---
title: "p1805_hw3_jc5457"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```



### Problem 2
#### Load and clean the data
```{r}
accel_df =
  read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(day = factor(day))
```

#### Describe the resulting dataset
There are 35 observations and 1443 variables in this dataset. 
The variables are `week`, `day_id`, `day`, and 1440 activity variables `activity_1`, `activity_2`... `activity_1440`. 

#### Traditional analyses of accelerometer data focus on the total activity over the day.
Here we use the tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. 
```{r}
accel_df %>% 
  mutate(total_activity = sum(accel_df[4:1443])) %>% 
  subset(select = c("week","day_id","day","total_activity" )) %>% 
  knitr::kable(digits = 2)
```

Trend: The total activity in each day through five weeks are all the same.

#### Making plot
Accelerometer data allows the inspection activity over the course of the day. Here we make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. 
```{r}
accel_df %>%
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity", 
    values_to = "activity_count_per_minute") %>%
  mutate(activity = factor(activity)
         ) %>% 
  ggplot(aes(x = day_id, 
             y = activity_count_per_minute, 
             colour = day_id)) +
  geom_point()
```

The graph shows that this person mainly had their activity count per minute below 2500.



### Problem 3
#### Load the data
```{r}
library(p8105.datasets)
data("ny_noaa")
```

#### Describe the data
This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns, describing the five core variables for all New York state weather stations from January 1, 1981 through December 31, 2010.

Key variables include:

* `id`: Weather station ID, with `r is.na(ny_noaa$id) %>% mean()*100`% of missing value.
* `date`: Date of observation, with `r is.na(ny_noaa$date) %>% mean()*100`% of missing value.
* `prcp`: Precipitation (tenths of mm), with `r is.na(ny_noaa$prcp) %>% mean()*100`% of missing value.
* `snow`: Snowfall (mm), with `r is.na(ny_noaa$snow) %>% mean()*100`% of missing value.
* `snwd`: Snow depth (mm), with `r is.na(ny_noaa$snwd) %>% mean()*100`% of missing value.
* `tmax`: Maximum temperature (tenths of degrees C), with `r is.na(ny_noaa$tmax) %>% mean()*100`% of missing value.
* `tmin`: Minimum temperature (tenths of degrees C), with `r is.na(ny_noaa$tmin) %>% mean()*100`% of missing value.


#### Do some data cleaning. 
Here we create separate variables for year, month, and day. 

Then, transfer the unit of observations: 

* temperature (0.1 degrees C to F), 
* precipitation (0.1 mm to inch),
* snowfall (mm to inch)

Then, visualize the data to find the most commonly observed values for snowfall.

```{r}
noaa_tidy = ny_noaa %>% 
  separate(date, sep="-", into = c("year","month","day")) %>% 
  mutate(year = as.numeric(year), 
         month = as.numeric(month), 
         day = as.numeric(day),
         prcp = as.integer(round(prcp*.1*0.03937008)),
         snow = as.integer(round(snow*0.03937008)),
         snwd = as.integer(round(snwd*0.03937008)),
         tmax = as.integer(round(as.numeric(tmax)*.1*1.8+32)),
         tmin = as.integer(round(as.numeric(tmin)*.1*1.8+32))
         )

noaa_tidy %>%
  ggplot(aes(x = snow)) + 
    geom_histogram(binwidth=1,na.rm=TRUE) +
    labs(title = "Figure Q3a. histogam for snowfall value frequency",
         x = "Snow (inches)",
         y = "Count")
```

From the plot we can see: For snowfall, the most commonly observed values are 0. Because snowfall usually takes place in some winter days in NY State, for most of the days in a year there's no snowfall.

#### Make a two-panel plot

Here we make a two-panel plot showing the average max `temperature` in January and in July in each station across `years`.
```{r}
noaa_avg = noaa_tidy %>% 
  filter((month == 1 | month == 7) & !is.na(tmax)) %>% 
  group_by(id,year,month) %>% 
  summarise(avg_tmax = mean(tmax), .groups = "keep") %>% 
  mutate(month = recode(month,`1` = "January", `7` = "July"),
         avg_tmax = round(avg_tmax))
  
ggplot(data = noaa_avg)+
  geom_line(aes(x = year,y = avg_tmax,color=id),
            show.legend = FALSE,alpha=.3
            )+
  geom_smooth(aes(x = year,y = avg_tmax),
              color="black",show.legend = FALSE
              )+
  facet_grid(~month)+
  labs(title = "Figure Q3b. two-panel plot of the average max temperature in January and in July in each station across years",
       x = "Year",
       y = "Average Max Temp (degrees F)"
       )+ 
  scale_x_continuous(n.breaks = 10)
```

1. observable / interpretable structure: 

* July's average max temperatures are higher than January's, which is expected as summer have higher temperatures than winter. 
* The average max temperatures in each station across the years fluctuate much more in January than that in July. 

2. Outliers:
  
* In January 1982 and July 1988 we can observe two outliers with much lower average temperatures comparing with other data points.


#### Make another two-panel plot
Here we make a two-panel plot showing (i) `tmax` vs `tmin` for the full dataset; and (ii) the distribution of `snowfall` values greater than 0 and less than 100 separately by year.

```{r}
t_vs = noaa_tidy %>% 
  pivot_longer(cols=c("tmax","tmin"),
               names_to = "t_cat",
               values_to = "t_val") %>% 
  ggplot(aes(x = t_val)) +
  geom_density(aes(fill = t_cat), 
               color = "green", alpha = .5,na.rm = TRUE)+
  labs(fill = '',
       title = "Figure Q3c. tmax vs tmin",
       x = "Temperature (Degrees F)",
       y = "Density")

snow_year = 
  ny_noaa %>% 
  select(id,date,snow) %>% 
  filter(snow>0 & snow<100) %>% 
  separate(date,sep="-",into=c("year","month","day")) %>% 
  mutate(
    year = as.factor(year),
    month = as.numeric(month),
    day = as.numeric(day)
    ) %>% 
  ggplot(aes(x = year,y = snow))+
  geom_violin(aes(fill = year),
              color="green",show.legend = FALSE)+
  theme(axis.text.x = element_text(angle = -75)) +
  labs(title = "Figure Q3d. distribution of snowfall (<100 mm) separated by year",
       x = "Year",
       y = "Snowfall (mm)")

t_vs/snow_year
```



